Diversification of Australian blind snakes

### 2022-12-19 manuscript.Rmd
update result secition biogeobears

Based on 50 BioGeoBEARS biogeographic stochastic mappings (BSM) of our best-fitting model. We estimated that an average of 71.2 ± 1.3 (mean ± SD) total anagenetic and cladogenetic events occurred throughout the phylogeny. Most of these events were anagenetic dispersals (31.2 ± 1.3 events) and cladogenetic founder events (6.1 ± 1.5 events).

@@

Based on current distributions (ALA) most _Anilios_ species are found entirely or at least partially in the tropical grassland biome (n = 20) followed by the arid biome (n = 13) and mediterranean biome (n = 10). The model indicates two pulses of biogeographic transitions for the two major clades, which we will refer to as *Group 1* (n=21 lineages) and *Group 2* (n=14 lineages) (\@ref(fig:fig-biogeobears)a. The common ancestor of *Group 1* remained mostly in the Tropical Grassland Biome, then 12/21 lineages expanded or transitioned into other biomes. The common ancestor of *Group 2* first transitioned into the arid biome, and then 7/14 lineages expanded to the Mediterranean and Temperate Forest biomes, and/or transitioned back into the Tropical Grassland Biome.

@@

With 47 recognised species, the mostly Australian-endemic genus *Anilios* is currently the most diverse genus of scolecophidia in the world. In this study we inferred the extant species in Australia has a tropical grassland origin. Within a few million years, this ancestral lineage split with one group staying in the tropical grassland biome while the other group transitioning into the arid biome. Subsequent transitions in these two groups resulted in the distribution of biomes we observed across the phylogeny today. We also found that paleo aridity index provides the best fit to diversification dynamics observed in our blind snake tree. Although aridification has been shown to accelerate diversification rates in other groups of vertebrates (CITE), we found no such difference for Australian *Anilios* blind snakes.

@@

Consistent with @marinTracing2013, we found the initial diversification occurred very rapidly (<1 million years) in the tropical grassland biome followed by transitions and/or expansion into other biomes. 
However, as opposedrecent invasion (<10 MYA) and diversification (<5 MYA) into the arid zone [@marinHidden2013], our results show that species in both groups 1 and 2 have deep divergence between 20 – 10 MYA, during the time when the Australia was still relatively wet and humid [@byrneBirth2008]. This suggests that these lineages have diversified before the intensive continent-wide aridification in the late Miocene that resulted in the present dry and sandy areas of the arid biome [@byrneBirth2008].

@@

@@

Our results suggest that range expansion was important for diversification. The fact that DEC+J+x model was preferred over DEC+J may be due to poor dispersal ability as closely related species occupy the same or nearby areas. Events like floods are likely to have faciliated dispersals but in short distances.

>>>>

Based on 50 BioGeoBEARS biogeographic stochastic mappings (BSM) of the DEC+_j_+_x_ model. We estimated that an average of 71.2 ± 1.3 (mean ± SD) total anagenetic and cladogenetic events occurred throughout the phylogeny (see supplementary_table_BioGEOBEARS). Of these 40 ± 0 were cladogenetic events compared to 31.2 ± 1.3 anagenetic dispersal events. Out of the 40 cladogenetic dispersal events, most were narrow sympatry range expansion events (23.8±2.1 events) followed by subset sympatry (5.2±2.3 events), vicariance (4.9 ± 1.3), and founder events (6.1 ± 1.5).

@@

Based on current distributions (ALA) most _Anilios_ species are found entirely or at least partially in the tropical grassland biome (n = 20) followed by the arid biome (n = 13) and mediterranean biome (n = 10). The model indicates two pulses of biogeographic transitions for each  of the two major clades, which we will refer to as *Group 1* (n=21 lineages) and *Group 2* (n=14 lineages) (\@ref(fig:fig-biogeobears)a. The common ancestor of *Group 1* remained mostly in the Tropical Grassland Biome, then 12/21 lineages expanded or transitioned into other biomes. The common ancestor of *Group 2* first transitioned into the arid biome, and then 7/14 lineages expanded to the Mediterranean and Temperate Forest biomes, and/or transitioned back into the Tropical Grassland Biome.

@@

With 47 recognised species, the mostly Australian-endemic genus *Anilios* is currently the most diverse genus of scolecophidia in the world. In this study we inferred the extant species in Australia has a tropical grassland origin. Within a few million years, this ancestral lineage split with one group staying in the tropical grassland biome while the other group transitioning into the arid biome. Subsequent transitions in these two groups resulted in the distribution of biomes we observed across the phylogeny today. We also found that paleo aridity index provides the best fit to diversification dynamics observed in our blind snake tree. Although aridification has been linked elevated diversification rates in other plants and animals [@brennanMass2017;@raboskyAnalysis2014;@hernandez-hernandezAridification2014;@jhaPhylogeny2021], we found no such association for Australian *Anilios* blind snakes.

@@

Consistent with @marinTracing2013, we found the initial diversification occurred very rapidly (<1 million years) in the tropical grassland biome followed by transitions and/or expansion into other biomes. 
However, our results show that species in both groups 1 and 2 have deep divergence between 20 – 10 MYA and gradually diversified during the time when the Australia was still relatively wet and humid [@byrneBirth2008] as opposed more recent invasion (<10 MYA) and diversification (<5 MYA) into the arid zone proposed by @marinHidden2013. This suggests that these lineages have diversified before the intensive continent-wide aridification in the late Miocene that resulted in the present dry and sandy areas of the arid biome [@byrneBirth2008].

@@

We found that closely related species share similar biome and distribution. 
Our results indicate that cladogenetic processes (sympatry, vicariance, and founder event) made up the majority of range expansion events. Negative values for the distance parameter _X_ suggest that range expansion is expected to decrease as distance increase (CITE BioGeoBEARs intepetration?).

@@

While there's not study to show, blind snakes probably have small ranges and poor dispersal ability as closely related species occupy the same or nearby areas. Stochastic events like floods may have facilitated dispersals but in short distances.
